---
title: "Parameter Uncertainty Propagation"
author: "Libby Mohr"
date: "12/3/2020"
output: html_document
---

Calculate isotopic enrichment associated with 10% increase in nitrate concentration.
```{r}
source("utils.R")
RstN <- 0.003678
AFadd <- 0.984
AFinit <- RstN/(1+RstN)

maxPercentCIncrease <- 0.1
maxAF <- (AFinit + maxPercentCIncrease*AFadd)/(1+maxPercentCIncrease)
maxDel <- AFtoDelta(maxAF, RstN)
```

Import Linx II data, calculate depth-normalized k values for assimilatory uptake and denitrification. 
```{r}
library(readxl)
library(tidyverse)

linx <- read_xlsx(path = "./LINXII.xlsx", 
                  sheet = "Data",
                  col_types = c(rep("text", times = 3), 
                                rep("numeric", times = 10)),
                  na = "NA")

racewayDepth <- 0.15 # meters
linx <- linx %>% 
  filter(!NO3 == 0) %>%
  mutate(ktot_norm = ktot * Velocity * Depth/racewayDepth,
         kden_norm = kden * Velocity * Depth/racewayDepth) %>% 
  mutate(ku_norm = ktot_norm - kden_norm) %>%
  mutate(ku_norm_log = log(ku_norm),
         kden_norm_log = log(kden_norm),
         ktot_norm_log = log(ktot_norm),
         NO3_log = log(NO3)) 

```

Use Linx II data to estimate parameters (slope, intercept, variance) for a simple linear model of log(ktot) vs. log(NO3) and a simple linear model of log(kden) vs. log(ktot).
```{r}
library(rjags)
library(coda)
library(eivtools)
# Make data into a list that can be passed to JAGS
data <- list(x = linx$NO3_log, kTot = linx$ktot_norm_log, kDen =linx$kden_norm_log, n = nrow(linx))

## specify priors
data$b0 <- as.vector(c(0,0))      ## regression beta means
data$Vb <- solve(diag(10000,2))   ## regression beta precisions
data$s1 <- 0.1                    ## error prior n/2
data$s2 <- 0.1  

# Define model code
multivariate_regression <- "
model{

  betakTot ~ dmnorm(b0,Vb)      ## multivariate Normal prior on vector of regression params
  betakDen ~ dmnorm(b0,Vb)      ## multivariate Normal prior on vector of regression params
  preckTot ~ dgamma(s1,s2)   ## prior precision
  preckDen ~ dgamma(s1,s2)   ## prior precision

  for(i in 1:n){
      mukTot[i] <- betakTot[1] + betakTot[2]*x[i]   
      kTot[i]  ~ dnorm(mukTot[i],preckTot)  
      mukDen[i] <- betakDen[1] + betakDen[2]*kTot[i]   
      kDen[i]  ~ dnorm(mukDen[i],preckDen)  
  }
}"

# Specify number of chains, initial conditions
nchain = 3
inits <- list()
for(i in 1:nchain){
 inits[[i]] <- list(betakTot = rnorm(2,0,5), betakDen = rnorm(2,0,5), preckTot = runif(1,1/100,1/20), preckDen = runif(1,1/100,1/20))
}

# Create JAGS model object
j.model   <- jags.model(file = textConnection(multivariate_regression),
                        data = data,
                        inits = inits,
                        n.chains = nchain)
# Sample posterior distribution 
var.out   <- coda.samples (model = j.model,
                           variable.names = c("betakTot", "betakDen", "preckTot", "preckDen"),
                           n.iter = 5000)
# Evaluate Model
plot(var.out)
gelman.diag(var.out)
GBR <- gelman.plot(var.out)

# Remove burn-in
burnin = 500                                
jags.burn <- window(var.out,start=burnin)  
plot(jags.burn)   
# Convert output to matrix
out <- as.matrix(jags.burn)

```

Test Monte Carlo code for a single delta value, NO3 concentration, and gas exchange rate constant.
```{r}
# Specify number of realizations, n
n <- 10
# Define indices to sample MCMC output
indices <- sample.int(nrow(out), n, replace = TRUE)
# Specify simulation time
simTime <- seq(0, 60*24, length.out = 500)

# Generate ensemble of predictions for delta N15-NGas vs. time
ensemble <- predictNGas(N15NO30 = 0.5, 
                        NO30 = 10,
                        deltaNO30 = AFtoDelta((AFinit + (1/20)*AFadd)/(1+(1/20)), RstN),
                        B0Tot = out[indices,"betakTot[1]"], 
                        B1Tot = out[indices,"betakTot[2]"], 
                        precTot = out[indices,"preckTot"], 
                        B0Den = out[indices, "betakDen[1]"],
                        B1Den = out[indices,"betakDen[2]"], 
                        precDen = out[indices,"preckDen"], 
                        k2 = 0.02, 
                        n = n, 
                        time = simTime)

# Calculate 95% prediction interval limits at each time
pi <- getPI(ensemble, simTime)
# Create data frame of delta values for plotting
delta <- data.frame(t(ensemble$delta))%>%
  mutate(time = simTime/60) %>%
  pivot_longer(cols = 1:n)
# Plot each of 10 realizations
ggplot(delta, aes(x = time, y = value, color = name)) + 
  geom_line()
# Plot prediction interval with median
ggplot(data = pi, aes(x = time)) +
  geom_ribbon(aes(x = time, ymin = Lower, ymax = Upper), fill = "grey80") + 
  geom_line(aes(y = Median)) +
  xlab("Time (hours)") +
  ylab("Delta 15N-NGas")+
  theme_minimal()
```

Run Monte Carlo for 3 initial nitrate concentrations, 3 nitrate enrichments, and 3 gas exchange constants. 
```{r}
# Specify number of realizations, n
n <- 1000
# Define indices to sample MCMC output
indices <- sample.int(nrow(out), n, replace = TRUE)
# Specify simulation time
simTime <- seq(0, 60*24, length.out = 500)

# Generate data frame of parameter combinations
NO30 <- c(10, 100, 200)
deltaNO30 <- c(5000, 10000, 20000)
N15NO30 <-  matrix(calcPercN15(deltaNO30),nrow = 3) %*% matrix(NO30, nrow = 1) # Columns are delta values, rows are NO3 concentrations
k2 <- c(0.01, 0.1, 1)
params <- data.frame(
  N15NO30 = rep(as.vector(N15NO30), times = length(k2)),
  NO30 = rep(rep(NO30, each = length(deltaNO30)), times = length(k2)), 
  deltaNO30 = rep(rep(deltaNO30, times = length(NO30)), times = length(k2)), 
  k2 = rep(k2, each = (length(as.vector(N15NO30))))
)

# For each combination of parameters, generate ensemble of predictions for delta 15N-NGas vs. time
ensembles <- pmap(params, 
                  predictNGas,
                  B0Tot = out[indices,"betakTot[1]"], 
                  B1Tot = out[indices,"betakTot[2]"], 
                  precTot = out[indices,"preckTot"], 
                  B0Den = out[indices, "betakDen[1]"],
                  B1Den = out[indices,"betakDen[2]"], 
                  precDen = out[indices,"preckDen"], 
                  n = n, 
                  time = simTime)

# Calculate median, 2.5% and 97.5% quantiles for each time
PI0.01 <- map_dfr(ensembles[which(params$k2 == 0.01)], getPI, time = simTime)
PI0.1 <- map_dfr(ensembles[which(params$k2 == 0.1)], getPI, time = simTime)
PI1 <- map_dfr(ensembles[which(params$k2 == 1)], getPI, time = simTime)

# Plot prediction intervals with median
p0.01 <- plotPI(PI0.01) + ggtitle(expression(paste("k2 = 0.01 min"^"-1")))
p0.1 <- plotPI(PI0.1) + ggtitle(expression(paste("k2 = 0.1 min"^"-1")))
p1 <- plotPI(PI1) + ggtitle(expression(paste("k2 = 1 min"^"-1")))

plotPI(PI0.01) + ggtitle(expression(paste("k2 = 0.01 min"^"-1")))
plotPI(PI0.1) + ggtitle(expression(paste("k2 = 0.1 min"^"-1")))
plotPI(PI1) + ggtitle(expression(paste("k2 = 1 min"^"-1")))

# Save plots
ggsave("plots/k0.01.png", plot = p0.01, device = png())
ggsave("plots/k0.1.png", plot = p0.1, device = png())
ggsave("plots/k1.png", plot = p1, device = png())


```
